---
title: "Data Analysis"
output: pdf_document
date: "2023-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(aplore3)
library(tidyverse)
library(ggmosaic)
library(ggrepel)
```

```{r}

DF <- nhanes

str(DF)

summary(DF)

```
Removed the 37 records that missing the weight since its just 0.5% of total observations

```{r}

DF_1 <- DF[!(is.na(DF$wt)), ]

summary(DF_1)

for (i in c(2:ncol(DF_1))) {
  if (is.factor(DF_1[[i]])) {
    print(ggplot(DF_1, aes_string(x = colnames(DF_1)[i])) + geom_bar() + 
            labs(title = paste("Bar plot of", colnames(DF_1)[i]), x = colnames(DF_1)[i], 
                 y = "Count"))
  } else if (is.numeric(DF_1[[i]])) {
    print(ggplot(DF_1, aes_string(x = colnames(DF_1)[i])) + geom_histogram() +
            labs(title = paste("Histogram of", colnames(DF_1)[i]), x = colnames(DF_1)[i],
                 y = "Count"))
  } else {
    print(paste(colnames(DF_1)[i], "is not a factor or numeric. Skipping."))
  }
}

for (i in c(2:(ncol(DF_1)-1))) {
  if (is.factor(DF_1[[i]])) {
    print(ggplot(DF_1, aes_string(x = colnames(DF_1)[i])) + geom_bar(aes(fill = obese)) + 
            labs(title = paste("Bar plot of", colnames(DF_1)[i]), x = colnames(DF_1)[i], 
                 y = "Count"))
  } else if (is.numeric(DF_1[[i]])) {
    print(ggplot(DF_1, aes_string(x = colnames(DF_1)[i])) + geom_histogram(aes(fill = obese)) + facet_grid(~ obese) +
            labs(title = paste("Histogram of", colnames(DF_1)[i]), x = colnames(DF_1)[i],
                 y = "Count"))
  } else {
    print(paste(colnames(DF_1)[i], "is not a factor or numeric. Skipping."))
  }
}

```

## Association between categorical variables with the obese data

1) Obese vs Gender
```{r}

table(DF_1$obese, DF_1$gender)
prop.table(table(DF_1$obese, DF_1$gender), margin = 1)
chisq.test(table(DF_1$obese, DF_1$gender))

```

2) Obese vs Marital Status
```{r}

table(DF_1$obese, DF_1$marstat)
prop.table(table(DF_1$obese, DF_1$marstat), margin = 1)
chisq.test(table(DF_1$obese, DF_1$marstat))

```
3) Obese vs PSU
```{r}

table(DF_1$obese, DF_1$psu)
prop.table(table(DF_1$obese, DF_1$psu), margin = 1)
chisq.test(table(DF_1$obese, DF_1$psu))

```



4) Obese vs strata

```{r}

table(DF_1$obese, DF_1$strata)
prop.table(table(DF_1$obese, DF_1$strata), margin = 1)
chisq.test(table(DF_1$obese, DF_1$strata))

```

5) Obese vs vigwork

```{r}

table(DF_1$obese, DF_1$vigwrk)
prop.table(table(DF_1$obese, DF_1$vigwrk), margin = 1)
chisq.test(table(DF_1$obese, DF_1$vigwrk))

```



6) Obese vs modwork

```{r}

table(DF_1$obese, DF_1$modwrk)
prop.table(table(DF_1$obese, DF_1$modwrk), margin = 1)
chisq.test(table(DF_1$obese, DF_1$modwrk))

```


7) Obese vs wlkbik

```{r}

table(DF_1$obese, DF_1$wlkbik)
prop.table(table(DF_1$obese, DF_1$wlkbik), margin = 1)
chisq.test(table(DF_1$obese, DF_1$wlkbik))

```


8) Obese vs vigre excerise

```{r}

table(DF_1$obese, DF_1$vigrecexr)
prop.table(table(DF_1$obese, DF_1$vigrecexr), margin = 1)
chisq.test(table(DF_1$obese, DF_1$vigrecexr))

```



9) Obese vs modern excerise

```{r}

table(DF_1$obese, DF_1$modrecexr)
prop.table(table(DF_1$obese, DF_1$modrecexr), margin = 1)
chisq.test(table(DF_1$obese, DF_1$modrecexr))

```
3) Parameter Estimation and Confidence Intervals

```{r}

table(DF_1$obese) %>% prop.table()

```
1. Categories the age variable and get the proporion of obesity wrt to each age group and gender

```{r}
DF_1$Age_Category <- cut(DF_1$age, breaks = c(0, 18, 35, 50, 65, Inf), labels = c('0 - 18','19 - 35','36 - 50','51 - 65','66+'))

DF_1$Obesity <- ifelse(DF_1$obese == "Yes", 1, 0)

Age_Gender_Obese <- DF_1 %>%
  group_by(Age_Category, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count)

# Print the result
Age_Gender_Obese

ggplot(Age_Gender_Obese, aes(x = Age_Category, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Age Category", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)

```


2. Categories the age variable and get the proportion of obesity wrt to each minutes of sedernity and gender

```{r}
DF_1$Sedmin_Category <- cut(DF_1$sedmin, breaks = c(0, 10, 30, 60, 120, 240,360, 720, 840), labels = c('0 - 10','11 - 30','31 - 60','61 - 120','121 - 240', '241 - 360', '361 - 720', '721 - 840'))

Sedmin_Gender_Obese <- DF_1 %>%
  group_by(Sedmin_Category, gender, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
Sedmin_Gender_Obese

ggplot(Sedmin_Gender_Obese, aes(x = Sedmin_Category, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid( gender  ~ Age_Category) +
  labs(x = "Minutes of Sedentary Activity per Week", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



Sedmin_Gender_Obese_1 <- DF_1 %>%
  group_by(Sedmin_Category, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
Sedmin_Gender_Obese_1

ggplot(Sedmin_Gender_Obese, aes(x = Sedmin_Category, y = proportion, fill = Age_Category)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ Age_Category) +
  labs(x = "Minutes of Sedentary Activity per Week", y = "Proportion of Obesity", fill = "Age_Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

Sedmin_Gender_Obese_2 <- DF_1 %>%
  group_by(Sedmin_Category, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
Sedmin_Gender_Obese_1

ggplot(Sedmin_Gender_Obese, aes(x = Sedmin_Category, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Minutes of Sedentary Activity per Week", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
3. Categories the age variable and get the proportion of obesity wrt to vigerous work and gender and age

```{r}

vigwrk_Obese <- DF_1 %>%
  group_by(vigwrk, gender, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()


ggplot(vigwrk_Obese, aes(x = vigwrk, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid( gender  ~ Age_Category) +
  labs(x = "Vigorous Work Activity", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)



vigwrk_Obese_1 <- DF_1 %>%
  group_by(vigwrk, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
vigwrk_Obese_1

ggplot(vigwrk_Obese_1, aes(x = vigwrk, y = proportion, fill = Age_Category)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ Age_Category) +
  labs(x = "Vigorous Work Activity", y = "Proportion of Obesity", fill = "Age_Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)

vigwrk_Obese_2 <- DF_1 %>%
  group_by(vigwrk, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
vigwrk_Obese_2

ggplot(vigwrk_Obese_2, aes(x = vigwrk, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Vigorous Work Activity", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)



```



4. Categories the age variable and get the proportion of obesity wrt to Moderate work and gender and age

```{r}

modwrk_Obese <- DF_1 %>%
  group_by(modwrk, gender, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

modwrk_Obese

ggplot(modwrk_Obese, aes(x = modwrk, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid( gender  ~ Age_Category) +
  labs(x = "Moderate Work Activity", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)

modwrk_Obese

modwrk_Obese_1 <- DF_1 %>%
  group_by(modwrk, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
modwrk_Obese_1

ggplot(modwrk_Obese_1, aes(x = modwrk, y = proportion, fill = Age_Category)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ Age_Category) +
  labs(x = "Moderate Work Activity", y = "Proportion of Obesity", fill = "Age_Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


modwrk_Obese_2 <- DF_1 %>%
  group_by(modwrk, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
modwrk_Obese_2

ggplot(modwrk_Obese_2, aes(x = modwrk, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Moderate Work Activity", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


```


5. Categories the age variable and get the proportion of obesity wrt to Walk or Bicycle and gender and age

```{r}

wlkbik_Obese <- DF_1 %>%
  group_by(wlkbik, gender, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

wlkbik_Obese

ggplot(wlkbik_Obese, aes(x = wlkbik, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid( gender  ~ Age_Category) +
  labs(x = "Walk or Bicycle", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


wlkbik_Obese_1 <- DF_1 %>%
  group_by(wlkbik, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
wlkbik_Obese_1

ggplot(wlkbik_Obese_1, aes(x = wlkbik, y = proportion, fill = Age_Category)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ Age_Category) +
  labs(x = "Walk or Bicycle", y = "Proportion of Obesity", fill = "Age_Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


wlkbik_Obese_2 <- DF_1 %>%
  group_by(wlkbik, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
wlkbik_Obese_2

ggplot(wlkbik_Obese_2, aes(x = wlkbik, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Walk or Bicycle", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


```

6. Categories the age variable and get the proportion of obesity wrt to Vigerous Exercise and gender and age

```{r}

vigwrk_Obese <- DF_1 %>%
  group_by(vigwrk, gender, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

vigwrk_Obese

ggplot(vigwrk_Obese, aes(x = vigwrk, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid( gender  ~ Age_Category) +
  labs(x = "Vigorous Recreational Activities", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


vigwrk_Obese_1 <- DF_1 %>%
  group_by(vigwrk, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
vigwrk_Obese_1

ggplot(vigwrk_Obese_1, aes(x = vigwrk, y = proportion, fill = Age_Category)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ Age_Category) +
  labs(x = "Vigorous Recreational Activities", y = "Proportion of Obesity", fill = "Age_Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


vigwrk_Obese_2 <- DF_1 %>%
  group_by(vigwrk, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
vigwrk_Obese_2

ggplot(vigwrk_Obese_2, aes(x = vigwrk, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Vigorous Recreational Activities", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


```



7. Categories the age variable and get the proportion of obesity wrt to Moderate Recreational Activities and gender and age

```{r}

modrecexr_Obese <- DF_1 %>%
  group_by(modrecexr, gender, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

modrecexr_Obese

ggplot(modrecexr_Obese, aes(x = modrecexr, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid( gender  ~ Age_Category) +
  labs(x = "Moderate Recreational Activities", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


modrecexr_Obese_1 <- DF_1 %>%
  group_by(modrecexr, Age_Category) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
modrecexr_Obese_1

ggplot(modrecexr_Obese_1, aes(x = modrecexr, y = proportion, fill = Age_Category)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ Age_Category) +
  labs(x = "Moderate Recreational Activities", y = "Proportion of Obesity", fill = "Age_Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


modrecexr_Obese_2 <- DF_1 %>%
  group_by(modrecexr, gender) %>%
  summarise(obesity_count = sum(Obesity), total_count = n(), 
            proportion = obesity_count / total_count) %>% na.omit()

# Print the result
modrecexr_Obese_2

ggplot(modrecexr_Obese_2, aes(x = modrecexr, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(~ gender) +
  labs(x = "Moderate Recreational Activities", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_text(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5)


```



## Confidence Intervals for the obesity percentages in each group

1) Obesity

i) Wald
```{r}
###################################################
### code chunk number 7: Inference.Rnw:682-690
###################################################

alpha <- 0.05
phat <- sum(DF_1$Obesity) / length(DF_1$Obesity)

#Wald

CI=c(phat-qnorm(alpha/2,lower=F)*sqrt(phat*(1-phat)/length(DF_1$Obesity)),
     phat+qnorm(alpha/2,lower=F)*sqrt(phat*(1-phat)/length(DF_1$Obesity)))
round(CI,3)
```

ii) Likelihood Ratio CI

```{r}
###################################################
### code chunk number 8: LRCI
###################################################
ll.binom=function(pi,n,y){
  phat=y/n
  return(2*(y*log(phat/pi)+(n-y)*log((1-phat)/(1-pi))))
}

ll.binom.scaled=function(pi,n,y,alpha=0.05,df=1){
  quant=qchisq(alpha,df,lower=F)
  return(ll.binom(pi,n,y)-quant)
}

curve(ll.binom(x,n=length(DF_1$Obesity),y=sum(DF_1$Obesity)),from=0.14,to=0.17, xlab="p",
      ylab=paste0("l(p)"))
abline(h=3.84,lty=2,col="red")


lower=uniroot(ll.binom.scaled,interval=c(0.1,0.15), 
              n=length(DF_1$Obesity),y=sum(DF_1$Obesity))
upper=uniroot(ll.binom.scaled,interval=c(0.16,0.165), 
              n=length(DF_1$Obesity),y=sum(DF_1$Obesity))
abline(v=c(lower$root,upper$root), lty=2,col="blue")
text(0.15,12,labels=paste0("CI=[",
                         round(lower$root,2),",",
                         round(upper$root,2),"]"))
```

iii)  Wilson
```{r}

y=sum(DF_1$Obesity)
n=length(DF_1$Obesity)
phat=y/n
qhat=1-phat
z=qnorm(0.025,lower=F)
  


numl=phat+z^2/(2*n) - z*sqrt(phat*qhat/n+z^2/(4*n^2))
numu=phat+z^2/(2*n) + z*sqrt(phat*qhat/n+z^2/(4*n^2))
denom=1+z^2/n
numl/denom
numu/denom
```

iv) Agresti-Coull

```{r}
#Agresti-Coull
tildep=(y+z^2/2)/(n+z^2)
tildep - z*sqrt(tildep*(1-tildep)/(n+z^2))
tildep + z*sqrt(tildep*(1-tildep)/(n+z^2))
```

v) Clopper Pearson
```{r}
# Clopper Pearson
qbeta(0.05/2,shape1=y,shape2=n-y+1)
qbeta(1-0.05/2,shape1=y+1,shape2=n-y)
```

v) Listing of 11 different confidence intervals
```{r}
# Listing of 11 different confidence intervals
library(binom)
binom.confint(y, n, conf.level = 0.95)



```
## CI for each two binary variables - Clopper Pearson

1) Gender

```{r}

Age_Gender_Obese$Low_CI <- qbeta(0.05/2,shape1=Age_Gender_Obese$obesity_count,
                                 shape2=Age_Gender_Obese$total_count - Age_Gender_Obese$obesity_count+1)

Age_Gender_Obese$High_CI <- qbeta(1-0.05/2,shape1=Age_Gender_Obese$obesity_count+1,
                                  shape2=Age_Gender_Obese$total_count - Age_Gender_Obese$obesity_count)


# Plot the data
ggplot(Age_Gender_Obese, aes(x = Age_Category, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Low_CI, ymax = High_CI), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Age Category", y = "Proportion of Obesity", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + geom_text_repel(aes(label=round(proportion, 2)), vjust=-0.3, position = position_dodge(0.9), size=3.5) + geom_text_repel(aes(label=paste0("[", round(Low_CI, 2), ", ", round(High_CI, 2), "]"), y=High_CI), vjust=-1.5, position=position_dodge(0.9), size=3.5, col="red")



```

